---
title: "figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(rnaturalearth)
library(stars)
library(patchwork)
library(extrafont)
```

# California and Mexico shapefiles:
```{r}
# Read this in from `rnaturalearth`
US <- ne_states(country = "United States of America", returnclass = "sf")
US_sf <- st_as_sf(US, coords = c("longitude", "latitude")) %>% 
        select(name)

mexico <- ne_countries(country = "Mexico", returnclass = "sf") %>% 
        select(name)

full <- rbind(US_sf, mexico)
full_crop <- st_crop(full, aoi)
full_t <- st_transform(full_crop, proj)



# Select for just the CA shapefile
cali <- full %>% 
  filter(name == "California")
```

# Kelp maxent:

## First panel: current kelp distribution
```{r}
kelp_pal <- RColorBrewer::brewer.pal(5, "BuGn")

kelp <- read_stars(here("maxent_runs", "kelp", "Kelp.asc"), crs = proj)

kelp_plot <- ggplot() + 
        theme_void() +
        coord_equal() +
        geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
        geom_stars(data = kelp) +
        coord_sf(xlim = c(-125,  -116.5),
                 ylim = c(31.4, 41.5)) +
        theme(panel.border = element_rect(color = "black", fill = NA),
              panel.background = element_rect(fill = "mintcream"),
              plot.title = element_text(face = "bold"),
              legend.position = "none") +
        labs(title = "Kelp, current",
             fill = "Suitability")
```

## Second panel: kelp distribution in 2050
```{r}
kelp_2050 <- read_stars(here("maxent_runs", "kelp", "Kelp_bo_4.5.asc"), crs = proj)

kelp_2050_plot <- ggplot() +
        theme_void() +
        theme_void() +
        coord_equal() +
        geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
        geom_stars(data = kelp_2050) +
        coord_sf(xlim = c(-125,  -116.5),
                 ylim = c(31.4, 41.5)) +
         theme(panel.border = element_rect(color = "black", fill = NA),
            panel.background = element_rect(fill = "mintcream"),
            plot.title = element_text(face = "bold"),
            text = (element_text(family = "Times New Roman"))
            ) +
         labs(title = "Kelp 2050, RCP 4.5",
              fill = "Suitability")

kelp_2050_plot

```
## Third panel: change in kelp suitability
```{r}
# Difference
kelp_diff <- kelp_2050 - kelp
# Plot
kelp_diff_plot <- ggplot() +
      theme_void() +
      theme_void() +
      coord_equal() +
      geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
      geom_stars(data = kelp_diff) +
      scale_fill_gradient2(low = "firebrick", mid = "floralwhite", high = "forestgreen", na.value = NA,
                           limits = c(-1, 1),
                           guide = guide_colorbar(title.position = "bottom")) +
      coord_sf(xlim = c(-125,  -116.5),
               ylim = c(31.4, 41.5)) +
      theme(panel.border = element_rect(color = "black", fill = NA),
            panel.background = element_rect(fill = "floralwhite"),
            plot.title = element_text(face = "bold"),
            text = (element_text(family = "Times New Roman"))
            ) +
      labs(title = "Kelp suitability change", fill = "Change")

kelp_diff_plot

```

### Final kelp fig
```{r}
# Combine
kelp_suits <- ((kelp_plot + theme(legend.position = "bottom")) + (kelp_2050_plot + theme(legend.position = "bottom")) &
        scale_fill_gradient(low = "mintcream", high = "darkgreen", na.value = NA,
                            breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                            limits = c(0, 1),
                            guide = guide_colorbar(title.position = "bottom")) &
        theme(plot.margin= unit(c(1,1,1,1), units = "mm"),
              legend.box.background = element_rect(fill = NA, color = "black"),
              legend.box.margin = margin(4,7,7,8),
              legend.title.align = 0.5
              )) + (kelp_diff_plot &
  theme(plot.margin = unit(c(1,1,1,1), units = "mm"),
        legend.box.background = element_rect(fill = NA, color = "black"),
        legend.box.margin = margin(4,7,7,8),
        legend.position = "bottom",
        legend.title.align = 0.5
  )) &
    theme(text = element_text(family = "Times New Roman"))

kelp_suits
# Save
ggsave(here("figs", "kelp_final.png"), plot = kelp_suits)

```

### First panel: current otter distribution
```{r}
# Otter current
otters <- read_stars(here("maxent_runs", "otters", "Enhydra_lutris.asc"), crs = proj)

otter_plot <- ggplot() + 
        theme_void() +
        coord_equal() +
        geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
        geom_stars(data = otters) +
        coord_sf(xlim = c(-125,  -116.5),
                 ylim = c(31.4, 41.5)) +
        theme(panel.border = element_rect(color = "black", fill = NA),
              panel.background = element_rect(fill = "mintcream"),
              plot.title = element_text(face = "bold"),
              legend.position = "none") +
        labs(title = "Otters, current",
             fill = "Suitability")


otter_plot
```
### Second panel: 2050 otter distribution
```{r}
# Otter 2050
otters_2050 <- read_stars(here("maxent_runs", "otters", "Enhydra_lutris_bo_4.5.asc"))
otter_2050_plot <- ggplot() + 
        theme_void() +
        coord_equal() +
        geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
        geom_stars(data = otters_2050) +
        coord_sf(xlim = c(-125,  -116.5),
                 ylim = c(31.4, 41.5)) +
        theme(panel.border = element_rect(color = "black", fill = NA),
              panel.background = element_rect(fill = "mintcream"),
              plot.title = element_text(face = "bold"),
              legend.position = "none") +
        labs(title = "Otters 2050, RCP 4.5",
             fill = "Suitability")


otter_2050_plot
```
### Third panel: otter suitability difference
```{r}
# Difference
ott_diff <- otters_2050-otters

otter_diff_plot <- ggplot() +
      theme_void() +
      theme_void() +
      coord_equal() +
      geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
      geom_stars(data = ott_diff) +
      scale_fill_gradient2(low = "firebrick", mid = "floralwhite", high = "forestgreen", na.value = NA,
                           limits = c(-1, 1),
                           guide = guide_colourbar(title.position = "bottom")) +
      coord_sf(xlim = c(-125,  -116.5),
               ylim = c(31.4, 41.5)) +
      theme(panel.border = element_rect(color = "black", fill = NA),
            panel.background = element_rect(fill = "floralwhite"),
            plot.title = element_text(face = "bold"),
            text = (element_text(family = "Times New Roman"))
            ) +
      labs(title = "Otter suitability change", fill = "Change")

otter_diff_plot

```
### Final otter figure
```{r}
otter_suits <- ((otter_plot & theme(legend.position = "bottom")) + (otter_2050_plot + theme(legend.position = "bottom")) &
        scale_fill_gradient2(low = "mintcream", mid = "orchid1", high = "palevioletred4", na.value = NA,
                            breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                            midpoint = 0.5,
                            limits = c(0, 1),
                            guide = guide_colourbar(title.position = "bottom")
                                                   ) &
        theme(plot.margin= unit(c(1,1,1,1), units = "mm"),
              legend.box.background = element_rect(fill = NA, color = "black"),
              legend.box.margin = margin(4,7,7,8),
              legend.title.align = 0.5
              )) + (otter_diff_plot &
  theme(plot.margin = unit(c(1,1,1,1), units = "mm"),
        legend.box.background = element_rect(fill = NA, color = "black"),
        legend.box.margin = margin(4,7,7,8),
        legend.position = "bottom",
        legend.title.align = 0.5)) &
    theme(text = element_text(family = "Times New Roman"))

otter_suits

ggsave(here("figs", "otter_final.png"), plot = otter_suits)
```

### First panel: proportional threats
```{r}
threats_prop <- read_stars(here("threats", "comb_threat_prop.tif"))

prop_threat_plot <- ggplot() + 
        theme_void() +
        coord_equal() +
        geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
        geom_stars(data = threats_prop) +
        coord_sf(xlim = c(-125,  -116.5),
                 ylim = c(31.4, 41.5)) +
        theme(panel.border = element_rect(color = "black", fill = NA),
              panel.background = element_rect(fill = "mintcream"),
              plot.title = element_text(face = "bold"),
              legend.position = "bottom",
              legend.title.align = 0.5) +
        labs(title = "Scaled threats", fill = "Threat level") +
  scale_fill_gradient2(low = "mintcream", mid = "yellow", high = "deeppink", na.value = NA,
                            breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                            midpoint = 0.5,
                            limits = c(0, 1),
                       guide = guide_colourbar(title.position = "bottom")
                                                   )

prop_threat_plot
```
### Second panel: high threat cells with otter suitability change
```{r}
threats <- raster(here("threats", "threat_oil.6.tif"))
threats_s <- read_stars(here("threats", "threat_oil.6.tif"))

threat_center <- st_as_sf(rasterToPoints(threats, fun = function(x){x==1}, spatial = TRUE), crs = proj)

threat_plot <- ggplot() +
      theme_void() +
      theme_void() +
      coord_equal() +
      geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
      geom_stars(data = ott_diff) +
      scale_fill_gradient2(low = "firebrick", mid = "mintcream", high = "forestgreen", na.value = NA,
                           limits = c(-1, 1),
                           guide = guide_colourbar(title.position = "bottom")) +
  
      geom_sf(data = threat_center, color = "darkblue", fill = NA, pch = 0, size = 1.2) +

      coord_sf(xlim = c(-123.5,  -118),
               ylim = c(33, 38)) +
      theme(panel.border = element_rect(color = "black", fill = NA),
            panel.background = element_rect(fill = "mintcream"),
            plot.title = element_text(face = "bold"),
            text = (element_text(family = "Times New Roman")),
            legend.position = "bottom",
            legend.title.align=0.5
            ) +
      labs(title = "High threat cells overlaid\nwith otter suitability change", fill = "Otter suitability change")

threat_plot
```
### Third panel: otter suitability with future otter distribution
```{r}
threat_plot_2050 <- ggplot() +
      theme_void() +
      theme_void() +
      coord_equal() +
      geom_sf(data = full_t, color = "lemonchiffon3", fill = "lavenderblush2", na.value = NA) +
      geom_stars(data = otters_2050) +
      scale_fill_gradient(low = "mintcream", high = "palevioletred4", na.value = NA,
                          guide = guide_colourbar(title.position = "bottom")
                                                   ) +
  
      geom_sf(data = threat_center, color = "darkblue", fill = NA, pch = 0, size = 1.2) +

      coord_sf(xlim = c(-123.5,  -118),
               ylim = c(33, 38)) +
      theme(panel.border = element_rect(color = "black", fill = NA),
            panel.background = element_rect(fill = "mintcream"),
            plot.title = element_text(face = "bold"),
            text = (element_text(family = "Times New Roman")),
            legend.position = "bottom",
            legend.title.align = 0.5
            
            ) +
      labs(title = "High threat cells overlaid\nwith otter habitat in 2050", fill = "2050 Otter suitability")

threat_plot_2050

threat_plot
```
### Combined threat figure
```{r}

otter_threats_final <- prop_threat_plot + threat_plot + threat_plot_2050 &
   theme(plot.margin= unit(c(1,1,1,1), units = "mm"),
              legend.box.background = element_rect(fill = NA, color = "black"),
              legend.box.margin = margin(4,7,7,8),
              legend.text = element_text(size = 10),
         text = element_text(family = "Times New Roman")
              ) &
    theme(text = element_text(family = "Times New Roman"))


otter_threats_final

ggsave(here("figs", "otter_threats_final.png"), otter_threats_final)

```

