---
title: "Otter Data Processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(raster)
library(sf)
library(sdmpredictors)
library(rnaturalearth)
library(ncdf4)
library(stars)
library(rNOMADS)
library(enmSdm)
library(stringr)

rasterOptions(maxmemory = 1e10)
```

## This is a document for processing raw data for the Conservation Planning final project.

```{r}
# Set base projection
proj <- crs("+proj=longlat +datum=WGS84 +no_defs")
```

## Global variables
California Extent:
```{r}
# Read this in from `rnaturalearth`
full <- ne_states(country = "United States of America", returnclass = "sf")
# Select for just the CA shapefile
cali <- full %>% 
  filter(name == "California")
# Pull an extent from this
cal_ext <- extent(cali)
# Edit CA extent to include more ocean
aoi <- extent(c(-125.5,
                -114.1191,
                31,
                42))
```

## Bio Oracle Layers:
```{r}
# make a list of datasets
datasets <- list_datasets(terrestrial = FALSE, marine = TRUE)
list_layers(datasets)
# Read in the ones we want
bo_layers <- load_layers(c("BO_ph", "BO_dissox", "BO_sstmax", "BO_sstmean", "BO_sstmin", "BO_bathymean", "BO_bathymin", "BO_bathymax"))
# Check plot to see
plot(bo_layers)
# Crop to the new extent
bo_crop <- crop(bo_layers, aoi)
# Test plot
plot(bo_crop)
# Save all these rasters
writeRaster(bo_crop$BO_ph, filename = here("maxent_inputs", "bo_ph.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_dissox, filename = here("maxent_inputs", "bo_dissox.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_sstmax, filename = here("maxent_inputs", "bo_sstmax.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_sstmean, filename = here("maxent_inputs", "bo_sstmean.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_sstmin, filename = here("maxent_inputs", "bo_sstmin.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_bathymean, filename = here("maxent_inputs", "bo_bathymean.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_bathymin, filename = here("maxent_inputs", "bo_bathymin.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bo_crop$BO_bathymax, filename = here("maxent_inputs", "bo_bathymax.asc"), format = "ascii", overwrite = TRUE)
```

```{r}
# Create a base raster from BioOracle
base <- raster(here("maxent_inputs", "bo_dissox.asc"))
```

## GBIF otter data

```{r}
# DOI for otter2: 10.15468/dl.eqcwg3

gbif_otter2 <- read_tsv(here("raw_data", "GBIF", "otter_obs_gbif.csv")) %>% 
  dplyr::select("day", "month", "year", "decimalLatitude", "decimalLongitude")

# Remove observations wtihout coordinates:
otter_nona <- gbif_otter2 %>% 
  drop_na("decimalLatitude", "decimalLongitude")
# Make it an sf object
otter_sf <- st_as_sf(otter_nona, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
# Test plot
plot(otter_sf)
# Save it
write_csv(otter_sf, here("scratch", "otter_obs_clean.csv"))
# Reformat for maxent
otter_maxent <- otter_sf %>% 
  mutate(species = "Enhydra lutris",
         longitude = sf::st_coordinates(.)[,1],
         latitude = sf::st_coordinates(.)[,2]) %>% 
  dplyr::select("species", "longitude", "latitude")

otter_maxent_df <- as_data_frame(otter_maxent) %>% 
  dplyr::select(!"geometry")
# Save
write_csv(otter_maxent_df, here("maxent_inputs", "otter_obs.csv"))
# Rewrite file 
otter_obs <- read_csv(here("maxent_inputs", "otter_obs.csv"))
# Convert to sf
otter_sf <- st_as_sf(otter_obs, coords = c("longitude", "latitude"))
# Crop out observations fr outside of CA (There's one)
otter_crop <- st_crop(otter_sf, extent(aoi))
# Save the cropped file
write_csv(otter_crop, here("maxent_inputs", "otter_obs_crop.csv"))
```

## Kelp biomass from Tom Bell
```{r}
# Read in CONUS Landsat grid:
lst <- st_read(here("raw_data", "CONUS_ARD_grid"), layer = "conus_ard_grid")
# Get the extent
lst_extent <- extent(lst)
# Make a raster by dividing this into 30m pixels
lst_raster <- raster(ext = lst_extent, res = 30, crs = crs(lst))
# Reproject to the bo_crop extent:
lst_reproject <- projectRaster(lst_raster, bo_crop)
# Crop to our AOI
ca_30m <- crop(lst_reproject, aoi)
```

```{r}
# Open the ncdf
kelp_2020 <- nc_open(here("raw_data", "kelpCanopyFromLandsat_2020_v2.nc"))
# Pull the variables into R
kelp_lat <- ncvar_get(kelp_2020, "lat")
kelp_lon <- ncvar_get(kelp_2020, "lon")
kelp_yr <- ncvar_get(kelp_2020, "year")
kelp_q <- ncvar_get(kelp_2020, "quarter")

tail(kelp_yr) # The last 4 columns are 2020
tail(kelp_q) # THe last 4 columns are quarters 1234, 1 = jan-march, 2 = apr-june, 3 = jul-sep, 4 = oct-dec
# Each season is a column!!!!!
kelp_biomass <- ncvar_get(kelp_2020, "biomass")

# Convert from arrays to vectors
kelp_latv <- c(kelp_lat)
kelp_lonv <- c(kelp_lon)
# Create a data frame
# Most recent data is summer and fall from 2019, winter and spring from 2020. Indexed accordingly..
kelp_df <- data.frame(lat = kelp_latv, lon = kelp_lonv, winter = kelp_biomass[, 145], spring = kelp_biomass[, 146], summer = kelp_biomass[,143], fall = kelp_biomass[,144])

unique(kelp_df$summer)
# Make it spatial
kelp_sf <- st_as_sf(kelp_df, coords = c("lon", "lat"))
st_crs(kelp_sf) <- (proj)
# plot(kelp_sf)
# Rasterize it
kelp_raster<- rasterize(kelp_sf, bo_crop, fun = "sum")

# Create a blank raster of our BO extent
blank <- bo_crop$BO_ph
blank_reclass <- reclassify(blank,
                            c(-Inf, Inf, 0))
plot(blank_reclass)

plot(kelp_raster)
# Then fill all the cells with missing values with zeroes
kelp_new <- cover(kelp_raster, blank_reclass)

plot(kelp_new)

plot(blank)
plot(kelp_raster$summer)
hist(kelp_new$summer)
unique(kelp_new$fall)
# Save em
writeRaster(kelp_new$winter, here("maxent_inputs", "kelp_janmar.asc"), format = "ascii", overwrite = TRUE)
writeRaster(kelp_new$spring, here("maxent_inputs", "kelp_aprjun.asc"), format = "ascii", overwrite = TRUE)
writeRaster(kelp_new$summer, here("maxent_inputs", "kelp_julsep.asc"), format = "ascii", overwrite = TRUE)
writeRaster(kelp_new$fall, here("maxent_inputs", "kelp_octdec.asc"), format = "ascii", overwrite = TRUE)
```

## Let's turn kelp biomass into observations:
```{r}
# Turn biomass estimates into observations:
# This functon samples from our probability distribution:
set.seed(500)
kelp_oct_count <- sampleRast(kelp_new$fall, 10000, replace = TRUE, prob = TRUE)
kelp_oct_ct_sf <- st_as_sf(as.data.frame(kelp_oct_count),
                           coords = c("x", "y"))
# Jitter the geometries to get rid of overlap:
kelp_oct_ct_jitter <- st_jitter(kelp_oct_ct_sf, factor = .0001)
plot(kelp_oct_ct_jitter)
# Format this for maxent
kelp_oct_maxent <- kelp_oct_ct_jitter %>% 
  mutate(species = "Kelp",
         longitude = sf::st_coordinates(.)[,1],
         latitude = sf::st_coordinates(.)[,2]) %>% 
  as.data.frame() %>% 
  dplyr::select("species", "longitude", "latitude")
# Save it:
write_csv(kelp_oct_maxent, here("maxent_inputs", "kelp", "kelp_oct_maxent.csv"))
```

## Kelp Observation data
from https://wildlife.ca.gov/Conservation/Marine/Kelp/Aerial-Kelp-Surveys
```{r}
# We need to turn these shapefiles into observation points for each kelp forest
# Read in data:
kelp_aerial <- read_sf(here("raw_data", "kelp", "BIO_CA_Kelp2016"), layer = "BIO_CA_Kelp2016", )
# Make an observation point at the center of each point
kelp_centroid <- st_centroid(kelp_aerial)
# Transform it to lat-lon, not degrees
kelp_t <- st_transform(kelp_centroid, proj)
# Save a Maxent ready data frame
kelp_observations <- kelp_t %>% 
  mutate(species = "Kelp") %>% 
  mutate(longitude = unlist(map(kelp_t$geometry,1))) %>% 
  mutate(latitude = unlist(map(kelp_t$geometry,2))) %>% 
  as.data.frame() %>% 
  dplyr::select(species, longitude, latitude)
# Save file:
write_csv(kelp_observations, here("maxent_inputs", "kelp", "kelp_observations.csv"))
```

# Additional BO environmental layers for kelp:

```{r}
bo_sal <- load_layers("BO_salinity")
bo_sal_crop <- crop(bo_sal, aoi)
writeRaster(bo_sal_crop$BO_salinity, here("maxent_inputs", "kelp", "bo_sal.asc"), format = "ascii", overwrite = TRUE)
```
## Wavewatch III data:
```{r}
# Generate links to download sites:
# Create a vector for the years and months
year <- c(2015:2018, by = 1)
month <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
# Paste all combos of them together
dates <- paste(rep(year, each = 12), month, sep = "")

links <- NULL
# Generate links!
for(date in dates){
  links[date] <- paste("ftp://polar.ncep.noaa.gov/pub/history/waves/multi_1/", date, "/gribs/multi_1.wc_10m.hs.", date, ".grb2", sep = "")
}

# Get these files!
for(link in links){
download.file(link, destfile = basename(link))
}
```

Grib files were converted into netcdf using the wgrib2 utility, available from https://www.cpc.ncep.noaa.gov/products/wesley/wgrib2/
and the following bash script:

```{bash}
# cd into current directory
cd Desktop/Bren/"2nd Year"/"270p Conservation Planning"/otters/raw_data/wavewatch
# use wgrib2 to convert to netcdf
for file in *.grb2; do wgrib2 "$file" -netcdf "${file/%ext/out}".nc; done
```

```{r}
# Generate a list of filenames..
files <- list.files(here("raw_data", "wavewatch"), pattern = ".nc")
# Average wave heights for each month
for(i in 1:length(files)){
  # Brick all the 3 hour readings for a month
  brick_t <- brick(here("raw_data", "wavewatch", files[i]))
  # Average them
  t_avg <- calc(brick_t, fun = mean)
  # Get the date as a name
  name <- str_extract(files[i], pattern = "20....")
  full_name <- paste(name, ".tif", sep = "")
  # Save them
  writeRaster(t_avg, here("scratch", "wavewatch", full_name), overwrite = TRUE)
}
```

```{r}
# Average these for the entire run
files <- list.files(here("scratch", "wavewatch"))
# Brick all these files
wave_b <- brick(here("scratch", "wavewatch", files))
# Average the whole brick
wave_avg <- calc(wave_b, mean)
#plot(wave_avg)
# Crop to aoi
wave_crop <- crop(wave_avg, aoi)
# Resample to the BioOracle cell size
wave_resample <- resample(wave_crop, bo_crop, method = "ngb")
plot(wave_resample)
# Save it
writeRaster(wave_resample, here("maxent_inputs", "kelp", "wave_ht.asc"), format = "ascii", overwrite = TRUE)
```

## Future Bio-Oracle layers

```{r}
#Try again with BO data from the internet
# This is for RCP 4.5 and 2050
# Get the layer names
bo_fut_names <- list.files(here("raw_data", "BO_4.5"))
# Stack them
bofuture_brick <- stack(here("raw_data", "BO_4.5", bo_fut_names))
crs(bofuture_brick) <- proj
# Crop them
bofuture_crop <- crop(bofuture_brick, aoi)
# Save them all
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Current.Velocity.Max.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_cvmax.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Current.Velocity.Mean.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_cvmean.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Current.Velocity.Min.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_cvmin.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Salinity.Max.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_salmax.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Salinity.Mean.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_sal.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Salinity.Min.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_salmin.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Temperature.Max.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_sstmax.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Temperature.Mean.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_sstmean.asc"), format = "ascii", overwrite = TRUE)
writeRaster(bofuture_crop$X2050AOGCM.RCP45.Surface.Temperature.Min.asc.BOv2_1, here("maxent_inputs", "kelp", "maxent_4.5", "bo_sstmin.asc"), format = "ascii", overwrite = TRUE)

```