---
title: "threat_layer_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(sf)
library(here)
library(rgdal)
library(doParallel)
library(rgeos)

rasterOptions(maxmemory = 1e12)
```

## This rMarkdown processes the layers for the threat analysis and creates a finalized threat layer.

```{r}
# Set base projection
proj <- crs("+proj=longlat +datum=WGS84 +no_defs")

aoi <- extent(c(-125.5,
                -114.1191,
                31,
                42))

# Read in blank raster with BO extent and resolution
base <- raster(here("maxent_inputs", "bo_dissox.asc"))
crs(base) <- proj
```

## Fishing pressure processing:

```{r}
# Read data
fish_press <- raster(here("raw_data", "All_Fisheries_Ton_Km2_1931_2005.tif"))
# Reproject the base raster to the fishing pressure crs (the following steps save memory)
base_reproj <- projectRaster(base, crs = crs(fish_press))
# Crop fishing to our AOI
fish_crop <- crop(fish_press, base_reproj)
# Resample to our 5 arcmin grid
fish_resample <- resample(fish_crop, base_reproj, fun = "sum")
# Reproject back into our normal CRS and resolution
fish_reproj <- projectRaster(fish_resample, base)
# Save
writeRaster(fish_reproj, here("threats", "fishing_pressure_unscaled.tif"))


fish_reproj <- raster(here("threats", "fishing_pressure_unscaled.tif"))
# Reclassify to a proportional weight for each cell
fish_prop <- fish_reproj/maxValue(fish_reproj)
```
# Vessel tracks processing

```{r}
# Get layer names
layers <- ogrListLayers(here("raw_data", "WestCoast.gdb"))

# Write a function that reads, projects, crops, and rasterizes each monthly layer
tracks_fx <- function(layer){
  library(raster)
  library(sf)
  library(here)
  library(rgeos)
  tracks <- st_read(here("raw_data", "WestCoast.gdb"), layers)
  tracks_t <- st_transform(tracks, proj)
  tracks_c <- st_crop(tracks_t, aoi)
  tracks_raster <- rasterize(tracks, base, fun = "count")
  name <- paste(layer, ".tif", sep = "")
  writeRaster(tracks_raster, here("scratch", name))
}

# Run it in parallel to speed up processing time
cl <- makeCluster(detectCores()-1)
registerDoParallel(cl)
clusterExport(cl, varlist = c("layers", "proj", "aoi", "base"))
parLapply(cl, layers, tracks_fx)
stopCluster(cl)

# Brick all these files:
# List them
track_rs <- list.files(here("scratch"), pattern = ".tif")
# Select just the tif's, not the XMLs
track_stack <- stack(here("scratch", track_rs[c(1,3,5,7,9,11,13,15,17,19,21,23)]))
# Make a new raster with the sum of all the cells
track_sum_2018 <- raster::calc(track_stack, fun = sum)
# Save it
writeRaster(track_sum_2018, here("threats", "tracks_sum_2018_unscaled.tif"), overwrite = TRUE)

# Reclassify to a proportional weight for each cell:
track_sum_2018 <- raster(here("threats", "tracks_sum_2018_unscaled.tif")) 
tracks_prop <- track_sum_2018/maxValue(track_sum_2018)
plot(tracks_prop)
```


# Oil leases and platforms
```{r}
# Download oil platforms
download.file("https://www.boem.gov/PC-plat.zip", destfile = "oil_platforms.zip")
# Download active oil leases
download.file("https://www.boem.gov/PC-lease.zip", destfile = "oil_leases.zip")

# Read in platform data
platforms_shp <- st_read(here("raw_data", "oil_platforms"), layer = "pc_plat")
# Reproject
platforms_proj <- st_transform(platforms_shp, proj)
# Crop
platforms_crop <- st_crop(platforms_shp, aoi)
# Reproject to a meters projection
plat_m <- st_transform(platforms_crop, 7801)
# Buffer by 500m
platforms_buffer <- st_buffer(plat_m, dist = 500)
# Reproject back into our normal projection
plat_proj_buff <- st_transform(platforms_buffer, proj)
# Rasterize
platforms_raster <- rasterize(platforms_crop, base, field = 1, background = 0)
#Save
writeRaster(platforms_raster, here("threats", "platforms.tif"), overwrite = TRUE)


# Read in lease data
lease_shp <- st_read(here("raw_data", "oil_leases"), layer = "BOEM_Pacific_Leases")
# Reproject
lease_proj <- st_transform(lease_shp, proj)
# Crop
lease_crop <- st_crop(lease_shp, aoi)
# Rasterize
lease_raster <- rasterize(lease_crop, base, field = 1, background = 0)
#Save
writeRaster(lease_raster, here("threats", "leases.tif"))

# Combine the two...
oil_comb <- lease_raster+platforms_raster
# Reclassify so that cells with oil = 1
oil_comb_r <- reclassify(oil_comb, c(-Inf, 0, 0, 0, Inf, 1))
# Save it
writeRaster(oil_comb_r, here("threats", "oil_combined.tif"))
oil_comb_r <- raster(here("threats", "oil_combined.tif"))
```



# Create combined threat layer:

With equal weights, it seems like oil is too heavily weighted:
```{r}
# Add the proportional layers together:
comb_threat <- tracks_prop + oil_comb_r + fish_prop
# Make proportional from 0-1:
comb_threat_prop <- comb_threat/maxValue(comb_threat)
# Save it
writeRaster(comb_threat_prop, here("threats", "comb_threat_prop.tif"))

# Make a raster for the top 25% of most threatened cells:
threat_top50 <- reclassify(comb_threat_prop, c(-Inf, 0.5, 0, 
                                               0.5, 1, 1))
writeRaster(threat_top50, here("threats", "threat_top50.tif"))
plot(threat_top50)
```
Oil seems to have too much weight in this scheme.
So we can diminish oil's value in the weighting scheme
```{r}
# Weight oil 0.75 of its original value:
threat2 <- tracks_prop + fish_prop + 0.6*(oil_comb_r)
# Rescale from 0-1:
threat2_prop <- threat2/maxValue(threat2)
# Then take the top 50$ of raster values
threat2_top50 <- reclassify(threat2_prop, c(-Inf, 0.5, 0, 
                                               0.5, 1, 1))


plot(threat2_top50)
plot(threat2_prop)

writeRaster(threat2_top50, here("threats", "threat_oil.6.tif"))
```

